---
title: "In Class Assignment"
author: "Kshama Tavva"
date: "2026-10-06"
output: html_document
---
```{r}
library(tidyverse)
library(srvyr)
library(gssr)
library(gssrdoc)
library(haven)
data(gss_all)
wt_vars <- c("vpsu",     # PSU
             "vstrat",   # Strata          
             "wtssps",   # Weights
             "wtssnrps"  #  Weights with non-response adjustment
             )   
gss_all<-gss_all|>
  zap_missing()|>
  mutate(across(wt_vars, .fns=~as.numeric(.x)))
head(gss_all)

```
### Question 2

Is Gen Z more conservative compared to previous generations?

If this trend exists, does it match trends in specific policy views or cultural beliefs, party ID, or overall assessments of specific politicians or groups?

What are some individual factors that have typically been associated with a more/less conservative lean among younger adults? Are there differences across demographic groups?

```{r}
library(srvyr)
#svyggs<-gss_all|>
 # filter(!is.na(vpsu))|>
 # filter(!is.na(cohort))|>
  #filter(!is.na(partyid))|>
  #filter(!is.na(polviews))|>
#  filter(!is.na(homosex))|>
 # as_survey_design(
  #  ids=vpsu,
   # strata=vstrat,
    #weights = wtssnrps,
  #  nest=TRUE
#  )
```

```{r}
vars<-c("year","age","cohort","sex", "race","polviews", "partyid",
        "attend","degree","wtssnrps","vstrat","vpsu")
gss<-gss_all|>
  select(any_of(vars))|>
  mutate(across(everything(), ~as.numeric(haven::zap_labels(.x))))
rm(gss_all);gc()
gss
```

```{r}
gss2<-gss|>
  mutate(partyid_recode=recode_values(partyid, 1~"Democrat",2~"Democrat",3~"Democrat",4~"Independent",5~"Republican",6~"Republican",7~"Other"))|>
  filter(!is.na(partyid))|>
  mutate(generation=case_when(
     cohort >= 1928 & cohort <= 1945 ~ "Silent Generation",
     cohort >= 1946 & cohort <= 1964 ~ "Baby Boomer",
     cohort >= 1965 & cohort <= 1980 ~ "Generation X",
     cohort >= 1981 & cohort <= 1996 ~ "Millennials",
     cohort >= 1997 & cohort <= 2012 ~ "Generation Z",
  ))
```


```{r}
library(ggplot2)
ggplot(data=gss2,mapping = aes(x=year,y=partyid_recode, color=generation ))+geom_line()+
  geom_point()+
  labs(title = "Party Identification Over Time")+
  xlab("Party_Identification")
```

