---
title: "Homework 5: Survey Data"
format:
  html:
    toc: true
    df-print: kable
    code-tools: true
    embed-resources: true
---

```{=html}
<style>
.qbox {
  background-color: #cccccc; 
  padding: 20px; 
  border: 1px solid #999999; 
  border-left: 10px solid teal; 
  border-radius: 10px;
  margin: 10px;
}

</style>
```

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

```

You'll use data from the 2024 round of the General Social Survey to answer these questions. Scroll to @sec-codebook at the end of this document for some additional details on the variables you'll need to complete this analysis.

# Data import

You can download GSS data [directly from their website](https://gss.norc.org/us/en/gss/get-the-data.html), but Kieran Healy [has created an R package](https://kjhealy.github.io/gssr/) that will automatically import `GSS` data into R and provides some additional helper functions for working with the data.

Download the `gssr` and `gssrdoc` packages by running the code below:

```{r, eval=FALSE}
install.packages('gssr', repos =
  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

install.packages('gssrdoc', repos =
  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

```

Then you'll want to run the code below to set up your data for analysis.

```{r}
library(tidyverse)
library(srvyr)
library(gssr)
library(gssrdoc)
library(haven)
gss<-gss_get_yr(year='2024')
wt_vars <- c("vpsu",     # PSU
             "vstrat",   # Strata          
             "wtssps",   # Weights
             "wtssnrps"  #  Weights with non-response adjustment
             )   

# removing user-missing values and converting the weights to numeric
gss<-gss|>
  zap_missing()|>
  mutate(across(wt_vars, .fns=~as.numeric(.x)))

```

```{r}

library(gssr)
library(dplyr)

data(gss_all, package = "gssr")   # loads the full object into memory

vars <- c("year", "age", "cohort", "sex", "race", "polviews", "partyid",
          "attend", "degree", "wtssall", "vstrat", "vpsu")

gss <- gss_all |>
  select(any_of(vars)) |>
  filter(year >= 2000) |>
  mutate(across(everything(), ~ as.numeric(haven::zap_labels(.x))))

rm(gss_all); gc()   # free the big object right away
table(gss$year)
```

```{r}
library(ggplot2)

conservative = case_when(gss$polviews >= 5 ~ 1, gss$polviews %in% 1:4 ~ 0)
# Percent conservative, by year and generation (ages 18-27)
yr <- svyby(~conservative, ~year + generation, young, svymean, na.rm = TRUE)

ggplot(yr, aes(x = year, y = conservative * 100, color = generation)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = (conservative - 1.96 * se) * 100,
                    ymax = (conservative + 1.96 * se) * 100),
                width = 0.5, alpha = 0.5) +
  labs(title = "Percent conservative among ages 18-27, by generation",
       x = "Survey year", y = "% conservative", color = NULL) +
  theme_minimal()
```
